\textsf{Are large language models superhuman chemists?}

### 2024-10-20 src/tex/appendix.tex

\clearpage
\subsection{Implementation}
An overview of the benchmarking pipeline implemented in \chembench is shown in \Cref{fig:process}. More detailed information can be found in the online documentation of the \chembench package at \url{https://lamalab-org.github.io/chem-bench/}.
\begin{figure}
    \centering
    \includegraphics[height=.8\textheight]{figures/process.pdf}
    \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}. \texttt{Prompter} objects are then created to handle different types of prompts for instruction-tuned and completion models.
    The \texttt{Task} Iterator differentiates between \gls{mcq} and other question types. For each task type, appropriate prompts are generated and passed to the \gls{llm}. The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models).
    The regex-based parsing is elaborate and also able to handle special cases such as scientific notation, or roman numerals.
    If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback. The parsed or extracted answers then undergo metric computation and refusal detection.}
    \label{fig:process}
\end{figure}

@@

@@

Our analysis of both log probabilities and prompting confidence reveals distinct calibration behaviors across different language models.
\GPTFourO demonstrates an overconfident tendency, often assigning high probabilities even to incorrect answers.
However, when \GPTFourO displays high confidence, it accurately predicts correct answers approximately 80\% of the time. In contrast, \LlamaThreeOneEightBInstruct confidence distribution is more evenly spread, with a majority of predictions centered around 0.5. High-confidence predictions from \LlamaThreeOneEightBInstruct are less frequent compared to \GPTFourO, and unlike \GPTFourO, high confidence does not necessarily correlate with a higher chance of correct answers.
\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figures/log_probs_calibration_plot_overall_filtered.pdf}
    \caption{\textbf{Reliability diagram of histogram of logit-based confidence estimates.} For this analysis we obtained the linear probability from the logprobs of the models. Only logprobs of the tokens corresponding to the answers were considered. Linear proabability was computed by taking exponential of logprobs (for sequences with multiple tokens, values were multiplied).  The plot shows the average predicted probability and the fraction of correct answers for each bin of linear probabilities. The ideal scenario is a diagonal line, indicating perfect calibration where the model's confidence aligns perfectly with the actual correctness. The \gls{ece} value quantifies the overall calibration performance, with a lower \gls{ece} indicating better calibration.}
    \label{fig:confidence_score_distributions}
    \script{plot_logprobs.py}
\end{figure}

@@

>>>>

\clearpage
\subsection{Implementation}
An overview of the benchmarking pipeline implemented in \chembench is shown in \Cref{fig:process}. More detailed information can be found in the online documentation of the \chembench package at \url{https://lamalab-org.github.io/chem-bench/}.
\begin{figure}
    \centering
    \includegraphics[height=.7\textheight]{figures/process.pdf}
    \caption{\textbf{Overview of the benchmarking pipeline implemented in \chembench.} The process begins with JSON files containing task data, which are used to create \texttt{Task} objects and initialize the \texttt{ChemBenchmark}. \texttt{Prompter} objects are then created to handle different types of prompts for instruction-tuned and completion models.
    The \texttt{Task} Iterator differentiates between \gls{mcq} and other question types. For each task type, appropriate prompts are generated and passed to the \gls{llm}. The responses are then processed using regex-based parsing methods specific to \gls{mcq} or numeric answers (after obtaining the relevant part of the response from the instruction-tuned models).
    The regex-based parsing is elaborate and also able to handle special cases such as scientific notation, or roman numerals.
    If the initial parsing is unsuccessful, the system employs an \gls{llm}-based extraction method as a fallback. The parsed or extracted answers then undergo metric computation and refusal detection.}
    \label{fig:process}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/human_tool_usage_by_topic.pdf}
    \script{human_tool_usage.py}
    \caption{\textbf{Tool usage by human scorers.} The plot shows the most commonly used tools by human participants.}
    \label{fig:tool_use}
\end{figure}

@@

Our analysis of both log probabilities and prompting confidence reveals distinct calibration behaviors across different language models.
\GPTFourO demonstrates an overconfident tendency, often assigning high probabilities even to incorrect answers.
However, when \GPTFourO displays high confidence, it accurately predicts correct answers approximately 80\% of the time. In contrast, \LlamaThreeOneEightBInstruct confidence distribution is more evenly spread, with a majority of predictions centered around 0.5. High-confidence predictions from \LlamaThreeOneEightBInstruct are less frequent compared to \GPTFourO, and unlike \GPTFourO, high confidence does not necessarily correlate with a higher chance of correct answers.

\begin{figure}[htb]
    \centering
    \includegraphics[width=\textwidth]{figures/log_probs_calibration_plot_overall_filtered.pdf}
    \caption{\textbf{Reliability diagram of histogram of logit-based confidence estimates.} For this analysis we obtained the linear probability from the logprobs of the models. Only logprobs of the tokens corresponding to the answers were considered. Linear proabability was computed by taking exponential of logprobs (for sequences with multiple tokens, values were multiplied).  The plot shows the average predicted probability and the fraction of correct answers for each bin of linear probabilities. The ideal scenario is a diagonal line, indicating perfect calibration where the model's confidence aligns perfectly with the actual correctness. The \gls{ece} value quantifies the overall calibration performance, with a lower \gls{ece} indicating better calibration.}
    \label{fig:confidence_score_distributions}
    \script{plot_logprobs.py}
\end{figure}

@@

\subsection{Question refusal and LLM fallback} \label{sec:refusal}

\paragraph{Refusal}

During the training of the \glspl{llm}, generally one of the steps is refusal training. 
Thus, in this stage of the training, the models are trained to prevent them of generating harmful, undesirable or violent outputs.
As a consequence, it can happen that the model refuses to answer some questions that they identify as potential adversarially crafted prompts.

To have a measure of how much refusal happens in \chembench, we count the number of refused questions by each model with a rule-based approach. 
This rule-based approach searches for the most common structures of the refusal completions made by \glspl{llm}.

\paragraph{LLM-Fallback}

As explained above, the parsing of the answers from the completions of the models are made by parsing an answer pattern imposed within the prompt.
For the questions in which \glspl{llm} do not follow this pattern to answer, we implement the so-called \gls{llm}-fallback which consists in passing the question and the completion to another model asking to extract the question or return "None" otherwise.
For the count, the model chosen for do this auxiliary extraction was Llama 3 70B accessed via the Groq API.
